---
title: 'HW3: Class 1 Survey Analysis'
author: "Byeongju Ryu"
date: "2024-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab: Class One Survey Analysis

### INSTRUCTIONS

Complete all tasks 1-7 using [**reproducible coding practices**]{.underline} (annotation of all code). Use R syntax to solve all problems (i.e. do not manually change values/variables in datasets or check them manually to make sure things are correct except as a double or triple check). Please cut and paste each problem before each R markdown chunk and annotate your code within the chunks. Compare and consolidate your answers. Select one member of your group to walk the class through how each problem was solved. Elect one member of your group to turn in your consolidated code with answers.

#### 1. Import class one survey data from our Github site (<https://github.com/kijohnson/ADA-2024>) (The dataset is called ‘Class 1 Survey Fall 2024_di.csv’), calling the R dataframe that you create C1survey.

```{r}
class_survey_csv <-
 read_csv(
   "https://raw.githubusercontent.com/kijohnson/ADA-2024/main/Class%201%20(Getting%20started%20and%20RR%20I)/Lab/Class%201%20Survey%20Fall%202024_di.csv"
   )
```

#### 2. Determine: a. the number of observations (i.e. the number of people who filled out the survey) and b. the number of variables in the dataframe.

```{r}
dim(class_survey_csv)
```

#### 3. a. Rename the column variables to something shorter and that is descriptive of what the variable is about (for example like_dogs for the ‘Do you like dogs?’ question variable) and b. Display that they are renamed.

##### Point to remember about renaming variables

```{r}
colnames(class_survey_csv)
```

```{r}
# rename variables in order of appearance
names(class_survey_csv)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan",
"fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby",
"program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length",
"top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state",
"city", "highest_educ_level")
```

#### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.

```{r}
sapply(class_survey_csv, class)
str(class_survey_csv)
summary(class_survey_csv)
```

#### 5. a. Using code, check the bday and bmonth variables for any unusual or missing values. If you find any, b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and c. after you have cleaned up any unusual values, find the median bday and month. \*Note: please make sure that your bday and month variables are numeric before doing statistics

```{r}
## Prep (Character -> Numeric)
# first check class again
class(class_survey_csv$bday)
class(class_survey_csv$bmonth)

# check missing/unusual values
class_survey_csv$bday
class_survey_csv$bmonth

# data cleaning


# convert bday variable to numeric using the as.numeric function for stats
class_survey_csv$bday_n <-as.numeric(class_survey_csv$bday)
class_survey_csv$bmonth_n <-as.numeric(class_survey_csv$bmonth)

# check the class of the new numeric bday and bmonth variable
class(class_survey_csv$bday_n)
class(class_survey_csv$bmonth_n)

## Steps
# get summary stats; the missing values (NA's)
summary(class_survey_csv$bday_n)
summary(class_survey_csv$bmonth_n)
table(class_survey_csv$bday_n, useNA = "always")
table(class_survey_csv$bday_n, useNA = "always")
```

#### 6. a.Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.

```{r}
##Step a. Create a new variable: bseason
class_survey_csv <- class_survey_csv %>%
 mutate(bseason = case_when(bmonth %in% c(12,1,2) ~ "Winter",
                            bmonth %in% c(3,4,5) ~ "Spring",
                            bmonth %in% c(6,7,8) ~ "Summer",
                            bmonth %in% c(9,10,11) ~ "Fall"))
class_survey_csv$bseason

##Step b. Print a table of seasons


##Step c. Sum the columns
```


#### 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #’s).

