---
title: 'HW3: Class 1 Survey Analysis'
author: "Byeongju Ryu"
date: "2024-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HW3: Class 1 Survey Analysis

### 0. Preparation
```{r}
# Install required packages if these are not installed
  # install.packages("haven") 
  # install.packages("readr") 
  # install.packages("dplyr")
  # install.packages("openxlsx")
  # install.packages("pacman")

# for SPSS (read_sav()), stata (read_dta()), SAS(read_sas()) and xpt (read_xpt()) files
library(haven) 

# for reading csv file (read_csv())
library(readr) 

# for data management
library(dplyr) 

# for exporting excel files
library(openxlsx) 

# we can also use the pacman package that automatically checks for packages, installs them if not found, and opens the associated library once installed
pacman::p_load(haven, readr, dplyr, openxlsx)
```

### 1. Import class one survey data from our Github site (<https://github.com/kijohnson/ADA-2024>) (The dataset is called ‘Class 1 Survey Fall 2024_di.csv’), calling the R dataframe that you create C1survey.

```{r}
# Read the C1 dataset from Github
class_survey_csv <- read_csv("https://raw.githubusercontent.com/kijohnson/ADA-2024/main/Class%201%20(Getting%20started%20and%20RR%20I)/Lab/Class%201%20Survey%20Fall%202024_di.csv")
```

### 2. Determine: a. the number of observations (i.e. the number of people who filled out the survey) and b. the number of variables in the dataframe.

```{r}
# 29 observations and 27 variables are included in the dataset.
dim(class_survey_csv)
str(class_survey_csv)
```

### 3. a. Rename the column variables to something shorter and that is descriptive of what the variable is about (for example like_dogs for the ‘Do you like dogs?’ question variable) and b. Display that they are renamed.

```{r}
# Display the current column names (variable names)
colnames(class_survey_csv)
```

```{r}
# Rename variables in order of appearance
names(class_survey_csv)[1:27]<-c("id", "like_cats", "like_dogs", "have_desert", "slogan",
"fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby",
"program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length",
"top_three","public_health_interest", "fav_num", "bday", "bmonth", "country", "state",
"city", "highest_educ_level")

# Display the renamed variables (columns)
colnames(class_survey_csv)
```

### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.

```{r}
sapply(class_survey_csv, class) # Solution 1 using sapply function
str(class_survey_csv) # Solution 2 using str function
table(sapply(class_survey_csv, class)) # Solution 3 is the best solution using table and sapply functions 23 character and 4 integer
summary(class_survey_csv) # Solution 4
```

### 5. a. Using code, check the bday and bmonth variables for any unusual or missing values. If you find any, b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and c. after you have cleaned up any unusual values, find the median bday and month. \*Note: please make sure that your bday and month variables are numeric before doing statistics

```{r}
# a.  Check the bday and bmonth varaibles
  ## Check the classes again
  class(class_survey_csv$bday)
  class(class_survey_csv$bmonth)

  ## Look at values to check missing/unusual values in bday and bmonth varriables
  class_survey_csv$bday
  class_survey_csv$bmonth

# b. Data cleaning
  ## Cleaning bday variable
  class(class_survey_csv$bday) # check that it is coded as character class
  class_survey_csv$bday<-ifelse(class_survey_csv$bday == "May 21-report 21", "21", class_survey_csv$bday)
  class_survey_csv$bday<-ifelse(class_survey_csv$bday == "March 31st", "31", class_survey_csv$bday)
  class_survey_csv$bday # check values 

  ## Cleaning bmonth variable
  class(class_survey_csv$bmonth)
  class_survey_csv$bmonth[class_survey_csv$bmonth=="May 21-report 5"]<-"5"
  class_survey_csv$bmonth[class_survey_csv$bmonth=="February"]<-"2"
  class_survey_csv$bmonth[class_survey_csv$bmonth=="September"]<-"9"
  class_survey_csv$bmonth[class_survey_csv$bmonth=="July"]<-"7"
  class_survey_csv$bmonth[class_survey_csv$bmonth=="March"]<-"3"
  class_survey_csv$bmonth # check values 

  ## Convert the bday and bmonth variables into numeric
    ### Convert bday variable  
    class_survey_csv$bday<-as.numeric(class_survey_csv$bday)
    table(class_survey_csv$bday, useNA = "always") # check values 
  
    ### Convert bmonth variable
    class_survey_csv$bmonth<-as.numeric(class_survey_csv$bmonth)
    table(class_survey_csv$bmonth, useNA = "always") # check values 
    
    ### Check the classes of bday and bmonth variables
    class(class_survey_csv$bday)
    class(class_survey_csv$bmonth)

# c. Calculate medians for bday and bmonth 
median(class_survey_csv$bday, na.rm=TRUE) # 14 (biased because of 0, which is not a value) How could we better design a survey to ensure this doesn't happen?
median(class_survey_csv$bmonth, na.rm=TRUE) # 7
```

### 6. a.Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.

```{r}
# a. Create a new variable: bseason (using dplyr)
class_survey_csv <- class_survey_csv %>% 
  mutate(bseason=case_when(bmonth %in% c(12,1,2) ~"Winter",
                           bmonth %in% c(3,4,5) ~"Spring",
                           bmonth %in% c(6,7,8) ~"Summer",
                           bmonth %in% c(9,10,11) ~"Fall"))
table(class_survey_csv$bmonth, class_survey_csv$bseason) # check variable

class_survey_csv$bseason_f<-factor(class_survey_csv$bseason, levels=c("Winter", "Spring", "Summer", "Fall"))

# b. Print a table of seasons
table(class_survey_csv$bmonth, class_survey_csv$bseason)
table(class_survey_csv$bmonth, class_survey_csv$bseason_f)

# c. Sum the columns
tab<-addmargins(table(class_survey_csv$bmonth, class_survey_csv$bseason, useNA = "always"), 1)
tab
tab<-addmargins(table(class_survey_csv$bmonth, class_survey_csv$bseason_f, useNA = "always"), 1)
tab

# Results: Births by season: Fall = 4 Winter = 9 Spring = 8 Summer = 8
```


### 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #’s).
```{r}
# Step 1. Define a question: Which statistical software is mostly preferred among the students in class?

# Step 2. Check the distribution of "stat_software" variable
table(class_survey_csv$stat_software)

# Results: Among 29 students, SPSS was the most preferred software among the students (n=10), followed by R (n=8), Excel (n=5), Stata (n=4), and SAS (n=1). One response indicating a preference for both R and SPSS was excluded.
```


